---
title: "ETC5521: Exploratory Data Analysis"
author: ["Emi Tanaka", "Di Cook"]
---

```{r setup, include = FALSE}
# please add any R-packages used here in DESCRIPTION imports
library(yaml)
library(tibble)
library(dplyr)
library(purrr)
library(glue)
library(knitr)
library(kableExtra)
knitr::opts_chunk$set(echo = FALSE)
```

```{r fns}
make_rectangular <- function(.data) {
  map_dfr(.data, ~tibble(!!!.x))
}


ui_list <- function(vec) {
  for(x in vec) {
    cat("-", x, "\n")
  }
}

```


```{r staffs}

dat <- read_yaml("_course.yml")
team_df <- make_rectangular(dat$team)
schedule_df <- make_rectangular(dat$schedule)

nlecturer <- sum(team_df$role=="lecturer")
ntutor <- sum(team_df$role=="tutor")
```


## Lecturer`r ifelse(nlecturer > 1, 's', '')`

```{r lecturers, results="asis"}
team_df %>% 
  filter(role=="lecturer") %>% 
  pull(name) %>% 
  ui_list()
```

## Tutor`r ifelse(ntutor > 1, 's', '')`

```{r tutors, results="asis"}
team_df %>% 
  filter(role=="tutor") %>% 
  pull(name) %>% 
  ui_list()
```

## Consultations

All consultations will be in using zoom. Check Moodle for the links.

```{r consult, results="asis"}
team_df %>% 
  mutate(consult_print = glue::glue("{paste0(name, ': ', consult)}")) %>% 
  pull(consult_print) %>% 
  ui_list()
```

## Expectations

* Unless your question is a personal issue or an administrative matter, please do not send us direct email. The moodle discussion forum is appropriate for questions related to the course.
* For personal or private administrative issues, the email contact is: ETC5521.Clayton-x@monash.edu
* Lectures are recorded, and you are expected to have either attended the lecture, or watched the recordings fully, prior to the tutorial for the week.
* Tutorials are not recorded, and attendance is expected.
* Bring your own laptop to the computer labs. If you do not have a laptop you can borrow one from MoVE
* If you register after the start of the semester or if you miss a lecture/tutorial, it is your responsibility to catch up with missed material, learn about due dates for material to be turned in, and getting assigned to a group for team work, as necessary.
* The computer software R and RStudio will be used for the unit.

## Tentative Schedule

* Lectures: **Mon 4-6pm** on zoom (see Moodle for link)
* Tutorials: **Mon 6-7.30pm** on zoom (see Moodle for link)

There are no lectures or tutorials during the midsemester break.

```{r}
schedule_df <- schedule_df %>% 
  mutate(HTML = NA, PDF = NA, tut = NA) %>% 
  mutate(slides = case_when(
    !is.na(HTML) & !is.na(PDF) ~ glue("<a class='fa fa-link' href='slides/{HTML}' target='_blank'></a> <a class='fa fa-file-pdf' href='slides/{PDF}' target='_blank'></a>"),
    !is.na(HTML) ~ glue("<a class='fa fa-link' href='slides/{HTML}' target='_blank'></a>"),
    !is.na(PDF) ~ glue("<a class='fa fa-file-pdf' href='slides/{PDF}' target='_blank'></a>"),
    TRUE ~ ""
  ),
  tut = case_when(
    !is.na(tut) ~ glue("<a class='fa fa-tasks' href='labs/{tut}.html' target='_blank'></a> <a class='fa fa-wrench' href='labs/{tut}.Rmd' target='_blank'></a> <a class='fa fa-lightbulb' href='labs/{tut}sol.html' target='_blank'></a> <a class='fa fa-laptop' href='labs/{tut}sol.Rmd' target='_blank'></a>"),
    TRUE ~ ""
  ),
  topic = ifelse(is.na(topic), "", topic),
  readings = ifelse(is.na(readings), "", readings),
  assessment = ifelse(is.na(assessment), "", assessment)) %>% 
  select(week, lecturer, slides, tut, topic, readings, assessment)


midsem_break <- 7 # after which week

schedule_df %>% 
  kable(escape = FALSE, align = c("c", "c", "c", "c", "l", "l"),
        col.names = c("Week", "Lecturer", "Slides", "Tutorial", 
                      "Topic", "Readings", "Assessments")) %>% 
  kable_styling(bootstrap_options = "striped",
                fixed_thead = T) %>% 
  column_spec(1, width = "0.5cm") %>% 
  column_spec(2, width = "1cm") %>% 
  column_spec(3, width = "1cm") %>% 
  group_rows("Midsemester Break (2 weeks)", midsem_break + 1, 12, label_row_css = "background-color: #d9e1e8;") %>% 
  add_indent(1:midsem_break) # this needs to be done to match up with above
```
