---
title: "Data Visualization with R <br> Workshop Day 2"
subtitle: Review of Grammar of Graphics
author: "Di Cook"
email: "dicook@monash.edu"
date: "12th Nov  2020 @ Statistical Society of Australia | Zoom"
length: "360 minutes"
color_theme: "yellow"
pdflink: "day2session0.pdf"
titleslide_bgimg: "images/d2bg3.jpg"
output:
  xaringan::moon_reader:
    css:
      - ninjutsu 
      - "assets/animate.css"
      - "assets/styles2.css"
      - "assets/fira-code.css"
      - "assets/monash-brand.css"
      - "assets/boxes.css"
      - "assets/custom.css"
    self_contained: false 
    seal: false 
    chakra: 'libs/remark-latest.min.js'
    lib_dir: libs
    mathjax: "assets/mathjax-local/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
    nature:
      highlightStyle: github
      highlightLanguage: r 
      highlightLines: true
      highlightSpans: false 
      countIncrementalSlides: false
      slideNumberFormat: '%current%/%total%'
      navigation:
        scroll: false 
        touch: true
        click: false
      ratio: '16:9'
---

```{r, include = FALSE}
current_file <- knitr::current_input()
```
```{r, include = FALSE, eval = F}
input <- fs::path_ext_set(current_file, "html")
pagedown::chrome_print(input = input, format = "pdf", wait = 25)
```
```{r titleslide, child="assets/titleslide2.Rmd"}
```
```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.path = "images/day2-session1/",
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  fig.retina = 3,
  warning = FALSE,
  message = FALSE,
  cache = FALSE
)
library(tidyverse)
filter <- dplyr::filter
select <- dplyr::select
```

---
class: font_smaller

# <img src="https://upload.wikimedia.org/wikipedia/commons/0/06/Who_logo.svg" width="50px" style="vertical-align: middle;"> Tuberculosis incidence

The TB data is from the [WHO]( https://www.who.int/tb/country/data/download/en/). 

.font_small[
```{r echo=FALSE}
tb <- tb <- read_csv(here::here("data/TB_notifications_2020-07-01.csv")) %>% 
  select(country, iso3, year, new_sp_m04:new_sp_fu) 
tb %>% 
  filter(country %in% c("Australia", "New Zealand")) %>% 
  DT::datatable(width=1150, height=100)
```
]

<div class="tag center animated rubberBand" style="position:absolute;top:2%;left:12%;transform:rotate(-3deg);">
demo data for today
</div>

---
class: myquestion middle

- Is the data in tidy form?
--

- What are the variables in this data?
--

- How many variables are there?
--

country (name, iso3), year, sex, age

---
class: center
#  <img src="https://upload.wikimedia.org/wikipedia/commons/0/06/Who_logo.svg" width="50px" style="vertical-align: middle;"> Tidying the data

.font_small[
```{r tbtidy, echo=FALSE}
tb <- read_csv(here::here("data/TB_notifications_2020-07-01.csv")) %>% 
  select(country, iso3, year, new_sp_m04:new_sp_fu) %>%
  pivot_longer(cols = new_sp_m04:new_sp_fu, names_to = "sexage", values_to = "count") %>%
  mutate(sexage = str_replace(sexage, "new_sp_", "")) %>%
  mutate(sex=substr(sexage, 1, 1), 
         age=substr(sexage, 2, length(sexage))) %>%
  filter(!(age %in% c("04", "014", "514", "u"))) %>% # duplication
  filter(year > 1996, year < 2013) %>% # missings and re-deefinition
  mutate(age_group = factor(age, # re-ecode
                            labels = c("15-24", "25-34", "35-44", 
                                       "45-54", "55-64", "65-"))) %>%
  select(country, year, age_group, sex, count)

# Select Australia
tb_oz <- tb %>% 
  filter(country == "Australia")

tb_oz %>% 
  DT::datatable(width = 1150, height = 100)
```
]

---

.font_small[
```{r tb_bar_prop, fig.show='hide'}
ggplot(tb_oz, aes(x = year, y = count, fill = sex)) +
  geom_bar(stat = "identity", position = "fill") + #<<
  facet_wrap(~age_group, ncol = 6) +
  scale_fill_brewer(name = "Sex", palette = "Dark2") +
  ylab("Proportion")
```

<br>
<br>
```{r ref.label='tb_bar_prop', echo=FALSE, fig.width=10, fig.height=3, out.width="90%"}
```
]
---
#  <img src="https://upload.wikimedia.org/wikipedia/commons/0/06/Who_logo.svg" width="50px" style="vertical-align: middle;"> What do we learn?

- Generally, proportion is higher for males
--

- Relative proportion is higher for older males 
--

- Quite variable proportions from year to year
--

<br>
<br>
<br>
What don't we learn from this plot?

```{r eval=FALSE, echo=FALSE}
Counts are completely lost
```

---

.font_small[
```{r tb_count_trend, fig.show='hide'}
ggplot(tb_oz, aes(x = year, y = count, colour = sex)) +
  geom_point() + #<<
  geom_smooth(se = F) + #<<
  facet_wrap(~age_group, ncol = 6) +
  scale_colour_brewer(name = "Sex", palette = "Dark2") 
```

<br>
<br>
```{r ref.label='tb_count_trend', echo=FALSE, fig.width=10, fig.height=3, out.width="90%"}
```
]

---
#  <img src="https://upload.wikimedia.org/wikipedia/commons/0/06/Who_logo.svg" width="50px" style="vertical-align: middle;"> What do we learn?

- Generally, counts are quite varied from year to year, but relatively stable
--

- Increasing trend in counts for both males and females under 35 
--

- Counts for males almost always higher than females

--

---

.font_small[
```{r tb_pie, fig.show='hide'}
tb_oz %>%
  filter(year == 2010) %>%
  ggplot(aes(x = 1, y = count, fill = sex)) +
    geom_bar(stat = "identity", position = "fill") + #<<
    facet_wrap(~age_group, ncol = 6) +
    scale_fill_brewer(name = "Sex", palette = "Dark2") +
    xlab("") + ylab("") +
    coord_polar(theta = "y") #<<
```

<br>
<br>
```{r ref.label='tb_pie', echo=FALSE, fig.width=10, fig.height=3, out.width="90%"}
```
]

---
#  <img src="https://upload.wikimedia.org/wikipedia/commons/0/06/Who_logo.svg" width="50px" style="vertical-align: middle;"> What do we learn?

In 2010,

- there were almost no 45-54 year old women with TB
--

- there were more 24-35 year old women with TB than men
--

- generally more males than females had TB

--

---
class: question middle center

How many plots should you usually do?
--

Lots


---


class: exercise middle hide-slide-number


<i class="fas fa-users"></i>

# <i class="fas fa-code"></i> Open `day2-exercise-01.Rmd`

<center>
`r countdown::countdown(15, class = "clock")`
</center>


---

```{r endslide, child="assets/endslide.Rmd"}
```

