---
title: "Data Visualization with R <br> Workshop Day 2"
subtitle: Making maps
author: "Di Cook"
email: "dicook@monash.edu"
date: "29th July 2020 @ Statistical Society of Australia | Zoom"
length: "360 minutes"
color_theme: "yellow"
pdflink: "day2-session2.pdf"
titleslide_bgimg: "images/d2bg5.jpg"
output:
  xaringan::moon_reader:
    css:
      - ninjutsu 
      - "assets/animate.css"
      - "assets/styles2.css"
      - "assets/fira-code.css"
      - "assets/monash-brand.css"
      - "assets/boxes.css"
      - "assets/custom.css"
    self_contained: false 
    seal: false 
    chakra: 'libs/remark-latest.min.js'
    lib_dir: libs
    mathjax: "assets/mathjax-local/MathJax.js?config=TeX-AMS-MML_HTMLorMML"
    nature:
      highlightStyle: github
      highlightLanguage: r 
      highlightLines: true
      highlightSpans: false 
      countIncrementalSlides: false
      slideNumberFormat: '%current%/%total%'
      navigation:
        scroll: false 
        touch: true
        click: false
      ratio: '16:9'
---

```{r, include = FALSE}
current_file <- knitr::current_input()
```
```{r, include = FALSE, eval = T}
input <- fs::path_ext_set(current_file, "html")
pagedown::chrome_print(input = input, format = "pdf", wait = 25)
```
```{r titleslide, child="assets/titleslide.Rmd"}
```
```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.path = "images/day2-session2/",
  fig.width = 6,
  fig.height = 4,
  fig.align = "center",
  out.width = "100%",
  fig.retina = 3,
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  cache = FALSE
)
```

```{r, include = FALSE}
library(tidyverse)
filter <- dplyr::filter
select <- dplyr::select
library(ggthemes)
library(viridis)
library(ggmap)
library(sf)
```

---
class: font_smaller

# <img src="https://upload.wikimedia.org/wikipedia/commons/0/06/Who_logo.svg" width="50px" style="vertical-align: middle;"> Tuberculosis incidence

The TB data is from the [WHO]( https://www.who.int/tb/country/data/download/en/). 

```{r tbtidy, echo=FALSE}
tb <- read_csv(here::here("data/TB_notifications_2020-07-01.csv")) %>% 
  dplyr::select(country, iso3, year, new_sp_m04:new_sp_fu) %>%
  pivot_longer(cols=new_sp_m04:new_sp_fu, names_to="sexage", values_to="count") %>%
  mutate(sexage = str_replace(sexage, "new_sp_", "")) %>%
  mutate(sex=substr(sexage, 1, 1), 
         age=substr(sexage, 2, length(sexage))) %>%
  dplyr::select(-sexage) %>%
  filter(!(age %in% c("04", "014", "514", "u"))) %>%
  filter(year > 1996, year < 2013) %>%
  mutate(age_group = factor(age, 
                            labels = c("15-24", "25-34", "35-44", 
                                       "45-54", "55-64", "65-"))) %>%
  dplyr::select(country, year, age_group, sex, count)
tb
```

---
class: question

What is a choropleth map?
--

<br>

Why use a choropleth map?
--

---
# How do we get a map?

A polygon map of the world can be extracted from the `maps` package. 

.font_small[
```{r worldmap}
world_map <- map_data("world")
world_map %>% 
  filter(region %in% c("Australia", "New Zealand")) %>% 
      DT::datatable(width=1150, height=100)
```
]

---
# Maps are basically groups of connected dots

.pull-left[
<br>
<br>

These are the points, defining the country boundary for Australia

.font_small[
```{r oz_points, fig.show='hide'}
oz <- world_map %>% 
  filter(region == "Australia")
ggplot(oz, aes(x = long, y = lat)) + 
  geom_point() + 
  coord_map() #<<
```
]
]

.pull-right[

<br>
<br>

```{r ref.label="oz_points", echo=FALSE, out.width="110%"}
```
]

---

.pull-left[
<br>
<br>
Connect the dots

.font_small[
```{r oz_lines, fig.show='hide'}
ggplot(oz, aes(x = long, y = lat, 
               group = group)) + 
  geom_point() + 
  geom_line() + #<<
  coord_map()
```
]

`r anicon::nia("What happened?", animate="tada")`
]

.pull-right[
<br>
<br>

```{r ref.label="oz_lines", echo=FALSE, out.width="110%"}
```
]
---

.pull-left[
<br>
<br>

Connect the dots

.font_small[
```{r oz_path, fig.show='hide'}
ggplot(oz, aes(x = long, y = lat, 
               group = group)) + 
  geom_point() + 
  geom_path() + #<<
  coord_map()
```
]
]

.pull-right[
<br>
<br>

```{r ref.label="oz_path", echo=FALSE, out.width="110%"}
```
]

---

.pull-left[
This map doesn't have states and territory connections.

.font_small[
```{r oz_state, fig.show='hide'}
ggplot(oz, aes(x = long, y = lat, 
               group = subregion)) + #<<
  geom_path() + 
  coord_map()
```
]
]

.pull-right[

```{r ref.label="oz_state", echo=FALSE, out.width="110%"}
```
]

---

.pull-left[
We can also plot the map using `geom_polygon`, and fill with colour.

.font_small[
```{r oz_polygon, fig.show='hide'}
ggplot(oz, aes(x = long, y = lat, 
               group = group)) + 
  geom_polygon() + #<<
  coord_map() 
```
]
]

.pull-right[

```{r ref.label="oz_polygon", echo=FALSE, out.width="110%"}
```
]

```{r eval=FALSE, echo=FALSE}
# Exercise 1, fill with a different colour
```

---

.pull-left[
Using a map theme makes the result look more map like

.font_small[
```{r oz_theme, fig.show='hide'}
ggplot(oz, aes(x = long, y = lat, 
               group = group)) + 
  geom_polygon() + 
  coord_map() +
  theme_map() #<<
```
]
]

.pull-right[

```{r ref.label="oz_theme", echo=FALSE, out.width="110%"}
```

]
---
class: transition middle animated slideInLeft

# Let's make a choropleth map of tuberculosis

---

Aggregate counts across sex and age group for 2012

.font_small[
```{r tb_map_gaps, fig.show='hide'}
tb_2012 <- tb %>%
  filter(year == 2012) %>%
  rename(region = country) %>%
  group_by(region) %>%
  summarise(count = sum(count))
ggplot(tb_2012, aes(map_id = region)) + 
    geom_map(aes(fill = count), map = world_map,  #<<
             color="grey70", size = 0.1, na.rm = TRUE) +  #<<
    expand_limits(x = world_map$long, y = world_map$lat) +
    scale_fill_viridis("Count") +
    theme_map() 
```
]

---

```{r ref.label='tb_map_gaps', echo=FALSE, out.width="100%", fig.width=12, fig.height=5.5}
```

---
class: question middle

.font_large[What happened to the USA? UK?]

---

# Check the name matching

.font_small[
```{r}
wm_names <- world_map %>% 
  select(region) %>% 
  distinct()
tb_names <- tb %>% 
  filter(year == 2012) %>% 
  select(country) %>% 
  distinct()
tb_miss_from_wm <- anti_join(tb_names, wm_names,
                             by=c("country" = "region")) #<<
DT::datatable(tb_miss_from_wm, width = 1150, height = 100)
```
]

---

.font_small[
```{r}
tb_fixed <- tb %>% 
  mutate(region=recode(country, 
                       "United States of America" = "USA", 
                       "United Kingdom of Great Britain and Northern Ireland" = "UK",
                       "Russian Federation" = "Russia",
                       "Viet Nam" = "Vietnam",
                       "Venezuela (Bolivarian Republic of)" = "Venezuela",
                       "Bolivia (Plurinational State of)" = "Bolivia",
                       "Czechia" = "Czech Republic",
                       "Iran (Islamic Republic of)" = "Iran",
                       "Iran (Islamic Republic of)" = "Laos",
                       "Democratic People's Republic of Korea" = "North Korea",
                       "Republic of Korea" = "South Korea",
                       "United Republic of Tanzania" = "Tanzania",
                       "Congo" = "Republic of Congo"))
```
]

---

`r emo::ji("smile")` `r anicon::nia("Try again!", animate="bounce")`

.font_small[
```{r tb_mapgeom, fig.show='hide'}
tb_2012 <- tb_fixed %>% #<<
  filter(year == 2012) %>%
  group_by(region) %>%
  summarise(count = sum(count))
ggplot(tb_2012, aes(map_id = region)) + 
    geom_map(aes(fill = count), map = world_map,  #<<
             color = "grey70", size = 0.1, na.rm = TRUE) +  #<<
    expand_limits(x = world_map$long, y = world_map$lat) +
    scale_fill_viridis("Count") +
    theme_map() 
```
]
---

```{r ref.label='tb_mapgeom', echo=FALSE, out.width="100%", fig.width=12, fig.height=5.5}
```

---
# Counts are typically skewed

.pull-left[

 may be best to symmetrise
 
.font_small[
```{r tb_map_counts, echo=TRUE, fig.show='hide'}
ggplot(tb_2012, aes(x = count)) + 
  geom_histogram() 
```
]
]

.pull-right[
```{r ref.label='tb_map_counts', echo=FALSE}
```
]

---

.pull-left[

.font_small[
```{r tb_map_logcounts, echo=TRUE, fig.show='hide'}
ggplot(tb_2012, aes(x = count)) + 
  geom_histogram() + 
  scale_x_log10()
```
]
]

.pull-right[
```{r ref.label='tb_map_logcounts', echo=FALSE}
```
]

---
# `geom_polygon` can also be used

.font_small[
```{r tb_map, fig.show='hide'}
tb_2012_map <- world_map %>% left_join(tb_2012) #<<
ggplot(tb_2012_map, aes(x = long, y = lat, group=group)) + 
    geom_polygon(aes(fill = count), #<<
             color="grey70", size = 0.1, na.rm = TRUE) +  #<<
    expand_limits(x = world_map$long*1.1, y = world_map$lat*1.1) +
    scale_fill_viridis("Count", trans = "log10") +
    theme_map() 
```
]

---

```{r ref.label='tb_map', echo=FALSE, out.width="100%", fig.width=12, fig.height=5.5}
```

```{r eval=FALSE, echo=FALSE}
# Exercise 2, use a square root scale
# Exercise 3, use a manual scale
```


---
# Resources

Where to go to get more help on maps

- `ozmaps` package: https://github.com/mdsumner/ozmaps, https://mdsumner.github.io/ozmaps/
- `eechidna` package https://docs.ropensci.org/eechidna/
- https://www.littlemissdata.com/blog/maps
- https://www.r-spatial.org/r/2018/10/25/ggplot2-sf.html
- https://www.paulamoraga.com/book-geospatial/sec-spatialdataandCRS.html

Note: The biggest change when converting spatial objects from a mapping software to a data analysis project is "thinning" the map to make it smaller and efficient to work with. 

---
class: transition middle animated slideInLeft

# Reading google maps and overlaying data

---

```{r eval=FALSE, echo=FALSE}
# Data extracted from ALA with this code
# install.packages("ALA4R")
library(ALA4R)
l <- specieslist("platypus")
platypus <- occurrences("Ornithorhynchus anatinus", 
                        email = "dicook@monash.edu", 
                        download_reason_id = 10)
# This now sends you an email to download from the website! And it is tsv not csv
platypus <- read_tsv(here::here("data_raw/platypus/data.csv"))
# There are some odd values!
oz_bbox <- c(112.9, # min long
              -45, # min lat
              159, # max long
              -10)
platypus <- platypus %>%
  filter(Longitude > oz_bbox[1], Longitude < oz_bbox[3], Latitude > oz_bbox[2], Latitude < oz_bbox[4])
save(platypus, file=here::here("data/platypus.rda"))
```

.pull-left[
.font_small[
```{r}
load(here::here("data/platypus.rda"))
p <- ggplot(platypus) + 
  geom_point(aes(x = Longitude, y = Latitude), 
             alpha = 0.1)
p
```
]

]
.pull-right[
<br>
<br>

.font_small[
```{r}
p + coord_map()
```
]
]

---
# Extract Open Street Map using `ggmap`

.font_small[
```{r eval=FALSE}
oz_bbox <- c(112.9, # min long
              -45, # min lat
              159, # max long
              -10) # max lat
oz_map <- get_map(location = oz_bbox, source = "osm")
save(oz_map, file=here::here("data/oz_map.rda"))
```
]

```{r eval=FALSE, echo=FALSE}
# Exercise 4, density plot
```

---

.pull-left[
.font_small[
```{r platypus, fig.show='hide'}
load(here::here("data/oz_map.rda"))
ggmap(oz_map) + 
  geom_point(data = platypus, 
             aes(x = Longitude, y = Latitude), 
             alpha = 0.1, colour = "orange") + 
  theme_map()
```
]
]
.pull-right[
```{r ref.label="platypus", echo=FALSE}
```

]
---

class: exercise middle hide-slide-number


<i class="fas fa-users"></i>

# <i class="fas fa-code"></i> Open `day2-exercise-02.Rmd`

<center>
`r countdown::countdown(15, class = "clock")`
</center>


---

```{r endslide, child="assets/endslide.Rmd"}
```

