---
title: "Part 2 Exercise 4: Determining the best plot design"
output:
  html_document:
    toc: true
    toc_float: true
---

Push the `knit` button!


```{r setup, child="exercise-setup.Rmd", eval = file.exists("exercise-setup.Rmd")}
```

```{r, include = FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  error = FALSE,
  warning = FALSE,
  fig.height = 3,
  fig.width = 10
)
```

```{r pkgs, message = FALSE, warning = FALSE}
library(tidyverse) # contains ggplot2, dplyr, tidyr, etc
```

## `tuberculosis` dataset

```{r tb-data}
tb <- read_csv(here::here("data/TB_notifications_2020-07-01.csv")) %>% 
  dplyr::select(country, iso3, year, new_sp_m04:new_sp_fu) %>%
  pivot_longer(cols=new_sp_m04:new_sp_fu, names_to="sexage", values_to="count") %>%
  mutate(sexage = str_replace(sexage, "new_sp_", "")) %>%
  mutate(sex=substr(sexage, 1, 1), 
         age=substr(sexage, 2, length(sexage))) %>%
  dplyr::select(-sexage) %>%
  filter(!(age %in% c("04", "014", "514", "u"))) %>%
  filter(year > 1996, year < 2013) %>%
  mutate(age_group = factor(age, 
                            labels = c("15-24", "25-34", "35-44", 
                                       "45-54", "55-64", "65-"))) %>%
  dplyr::select(country, year, age_group, sex, count)

# Filter Australia
tb_oz <- tb %>% 
  filter(country == "Australia") 
```

### Exercise 4.1: Which plot arrangement better reveals the spike in counts?

```{r part2-exercise-01, echo = F, fig.show='hold'}
ggplot(tb_oz, aes(x=year, y=count, fill=sex)) +
  geom_bar(stat="identity") + 
  facet_wrap(~age_group, ncol=6) +
  scale_fill_brewer("Sex", palette="Dark2") 
ggplot(tb_oz, aes(x=year, y=count, fill=sex)) +
  geom_bar(stat="identity", position="dodge") + 
  facet_wrap(~age_group, ncol=6) +
  scale_fill_brewer("Sex", palette="Dark2") 
```

```{r part2-exercise-04-01s, eval=F}
# add your answer here!
```

### Exercise 4.2: Rearrange this plot to better answer

"Is the proportion of TB incidence in males relative to females increasing with age?"

```{r part2-exercise-02, echo = T, fig.height = 4}
tb_oz %>% 
  filter(year == 2012) %>%
  ggplot(aes(x=1, y=count, fill=age_group)) +
  geom_bar(stat="identity", position="fill") + 
  facet_wrap(~sex, ncol=6) +
  scale_fill_brewer("", palette="Dark2") +
  xlab("") + ylab("") +
  coord_polar(theta = "y")
```

```{r part2-exercise-04-02s}
# add your code here!
```

### Exercise 4.3: This plot is missing something

By using proportions by sex, this plot lost some trend difference between the sexes over years. Fix it.

```{r part2-exercise-3, echo = T}
tb_oz %>% group_by(year, age_group) %>% 
  summarise(p = count[sex=="m"]/sum(count)) %>%
  ggplot(aes(x=year, y=p)) +
    geom_hline(yintercept = 0.50, colour="white", size=2) +
    geom_point() +
    geom_smooth(se=F) +
    facet_wrap(~age_group, ncol=6) +
    ylab("proportion of males")
```


```{r part2-exercise-04-03s}
# add your code here!
```
